---
title: "P8131 HW7 yz4184"
author: "Yunlin Zhou"
output: pdf_document
---

```{r, echo = T, message = FALSE, results='hide', warning=FALSE}

library(tidyverse)
library(nlme)
library(ggplot2)
library (lattice)
library(patchwork)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


```{r, message = FALSE, results='hide', warning=FALSE}
# import data
pitch_df = read.csv("./HW7-politeness_data.csv")%>%
  drop_na()%>%
  janitor::clean_names()
```

# (a) Exploratory analysis
## provide boxplots to show the relation between gender/attitude and pitch (ignoring different scenarios).

```{r fig.width = 8}
a.p1 = pitch_df %>% 
  ggplot(aes(x = gender, y = frequency)) +
  geom_boxplot()+
  labs(title = "relationship between gender and pitch")

a.p2 = pitch_df %>% 
  ggplot(aes(x = attitude, y = frequency)) +
  geom_boxplot()+
  labs(title = "relationship between attitude and pitch")

(a.p1 + a.p2)
```

\newpage

# (b)
## Fit a mixed effects model with random intercepts for different subjects (gender and attitude being the fixed effects).

```{r}
LMM1 <- lme (frequency ~ gender + attitude, random = ~1 | subject,  data = pitch_df, method='REML') 
summary (LMM1)
```


## What is the covariance matrix for a subject Yi?

## What is the covariance matrix for the estimates of fixed effects?

## What are the BLUPs for subject-specific intercepts?

## What are the residuals?

# (c)
## Fit a mixed effects model with intercepts for different subjects (gender, attitude and their interaction being the fixed effects).

```{r}
LMM2 <- lme (frequency ~ gender * attitude, random = ~1 | subject,  data = pitch_df, method='REML') 
summary (LMM2)
```

## Use likelihood ratio test to compare this model with the model in part (b) to determine whether the interaction term is significantly associated with pitch.

```{r}
LMM.1 <- lme (frequency ~ gender + attitude, random = ~1 | subject,  data = pitch_df, method='ML')

LMM.2 <- lme (frequency ~ gender * attitude, random = ~1 | subject,  data = pitch_df, method='ML')

anova(LMM.2,LMM.1) 
```


# (d)
## Write out the mixed effects model with random intercepts for both subjects and scenarios (gender and attitude being the fixed effects).

##  Fit the model using lmer in the lme4 package.

##  Write out the covariance matrix for a subject Yi.

## What is the interpretation of the coefficient for the fixed effect term attitude?





