---
title: "P8131 HW7 yz4184"
author: "Yunlin Zhou"
output:
  word_document: default
  pdf_document: default
---

```{r, echo = T, message = FALSE, results='hide', warning=FALSE}

library(tidyverse)
library(nlme)
library(ggplot2)
library (lattice)
library(patchwork)
library(lme4)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
```

```{r, message = FALSE, results='hide', warning=FALSE}
# import data
pitch_df = read.csv("./HW7-politeness_data.csv")%>%
  drop_na()%>%
  janitor::clean_names()
```

# (a) Exploratory analysis
## provide boxplots to show the relation between gender/attitude and pitch (ignoring different scenarios).

```{r fig.width = 8}
a.p1 = pitch_df %>% 
  ggplot(aes(x = gender, y = frequency)) +
  geom_boxplot()+
  labs(title = "relationship between gender and pitch")

a.p2 = pitch_df %>% 
  ggplot(aes(x = attitude, y = frequency)) +
  geom_boxplot()+
  labs(title = "relationship between attitude and pitch")

(a.p1 + a.p2)
```

As the plot shows above, we can conclude that female and informal attitude are tending to have higher frequency.

\newpage

# (b)
## Fit a mixed effects model with random intercepts for different subjects (gender and attitude being the fixed effects).

```{r}
LMM1 <- lme (frequency ~ gender + attitude, random = ~1 | subject,  data = pitch_df, method='REML') 
summary (LMM1)
```


## What is the covariance matrix for a subject Yi?

```{r}
VarCorr(LMM1)
VarCorr(LMM1)[1]
VarCorr(LMM1)[2]
```

sigma_b^2 = 598.1953

sigma^2 = 847.7049

sigma_b^2 + sigma^2 = 598.1953 + 847.7049 = 1445.9

For every subject, we have 14 measurements, so the covariance matrix is a 14*14 matrix.








## What is the covariance matrix for the estimates of fixed effects?

```{r}
vcov(LMM1)
```


## What are the BLUPs for subject-specific intercepts?

```{r}
random.effects(LMM1)
```


## What are the residuals?

```{r}
pitch_df$frequency-fitted(LMM1)
```

\newpage

# (c)
## Fit a mixed effects model with intercepts for different subjects (gender, attitude and their interaction being the fixed effects).

```{r}
LMM2 <- lme (frequency ~ gender + attitude + gender * attitude, random = ~1 | subject,  data = pitch_df, method='REML') 
summary (LMM2)
```


## Use likelihood ratio test to compare this model with the model in part (b) to determine whether the interaction term is significantly associated with pitch.

```{r}
LMM.1 <- lme (frequency ~ gender + attitude, random = ~1 | subject,  data = pitch_df, method='ML')

LMM.2 <- lme (frequency ~ gender + attitude + gender * attitude, random = ~1 | subject,  data = pitch_df, method='ML')

anova(LMM.2,LMM.1) 
```

Since the p-value for Likelihood ratio test is 0.2392 > 0.05, we fail to reject the null hypothesis and conclude that the interaction term is not significantly associated with pitch, at the significance level of 0.05.

\newpage

# (d)
## Write out the mixed effects model with random intercepts for both subjects and scenarios (gender and attitude being the fixed effects).




##  Fit the model using lmer in the lme4 package.

```{r}
LMM3 = lmer(frequency ~ gender + attitude + (1 | subject) +  (1 | scenario), data = pitch_df)
summary(LMM3)
```

##  Write out the covariance matrix for a subject Yi.

```{r}
VarCorr(LMM3)
VarCorr(LMM3)[1]
VarCorr(LMM3)[2]
```

sigma_b^2  = 224.4994 + 613.1903 = 837.6897

sigma^2 = 25.254^2 = 637.7645

sigma_b^2 + sigma^2 = 837.6897 + 637.7645 = 1475.454












## What is the interpretation of the coefficient for the fixed effect term attitude?

```{r}
fixed.effects(LMM3)
```

When the gender is fixed, the mean frequency of polite attitude will be 20.00238 units lower than informal attitude on average.


